The K2 Halo Photometry Campaign

### 2019-03-12 paper/ms.tex
gadigal
additions

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}
While the \kepler\ mission was designed to look at tens of thousands of faint stars ($V \gtrsim 12$), brighter stars which saturate the detector are nevertheless some of the most interesting because of the ease with which they can be observed by other instruments and the wealth of knowledge about them that is already available. By considering the unsaturated scattered light `halo' around these stars we retrieve precise light curves of most of the brightest stars in \ktwo fields from Campaign 6 onwards. This halo campaign reveals stellar variability ubiquitously, including effects of stellar pulsation, rotation, and binarity. Here we describe our pipeline, and present a catalogue of the halo sources, with classifications and parametrizations of their variability and remarks on interesting objects. These light curves are publicly available as a High Level Science Product. \href{https://github.com/benjaminpope/k2halo}{\color{linkcolor}\faGithub} % shamelessly borrowed from Luger!

@@

% describe saturation, smear

% describe TV-min algorithm

@@

% incorporation of k2sc

@@

% short cadence availability 
Seven stars in Campaign~13 and one in Campaign~16 were not only assigned long-cadence halo apertures, but also downloaded at short cadence. TV-min halo photometry appears to perform poorly on short-cadence data, however, and we do not consider the data products for these stars to be satisfactory for an initial data release. This may be to do with the total variation being a poor proxy for noise in the light curves, with relatively fewer thruster-firing jumps and relatively higher white noise. The long-cadence data appear to be of normal quality.

@@

% do k2 smear after that
The sample of K2 bright stars presented here only includes those with halo apertures, but while some others are available conventionally, many were not assigned target pixels and were not downloaded at all. Light curves for many of these stars can nevertheless be obtained with smear photometry \citep{Pope2016}, which uses collateral `smear' data to obtain a 1-D spatial profile with $\sim 1/1000$ of the flux on each CCD. These can be used to recover the 102 brightest otherwise-unobserved stars in nominal \kepler (Pope et al., in prep.), and also in K2, although in the latter case the sample is much smaller due to competition with halo apertures and the systematics correction is more difficult. A natural extension of both pieces of work would be to produce smear light curves of all bright stars without halo apertures in K2, which would finally make the \kepler extended mission magnitude-complete at the bright end.

% tess halo

@@

BJSP acknowledges being on the traditional territory of the Lenape Nations and recognizes that Manhattan continues to be the home to many Algonkian peoples. We give blessings and thanks to the Lenape people and Lenape Nations in recognition that we are carrying out this work on their indigenous homelands.
%

>>>>

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}
While the \kepler\ mission was designed to look at tens of thousands of faint stars ($V \gtrsim 12$), brighter stars which saturate the detector are nevertheless some of the most interesting because of the ease with which they can be observed by other instruments and the wealth of knowledge about them that is already available. By considering the unsaturated scattered light `halo' around these stars we retrieve precise light curves of most of the brightest stars in \ktwo fields from Campaign 6 onwards. This halo campaign reveals stellar variability ubiquitously, including effects of stellar pulsation, rotation, and binarity. Here we describe our pipeline, and present a catalogue of the halo sources, with classifications and parametrizations of their variability and remarks on interesting objects. These light curves are publicly available as a High Level Science Product from the Mikulski Archive for Space Telescopes (MAST). \href{https://github.com/benjaminpope/k2halo}{\color{linkcolor}\faGithub} % shamelessly borrowed from Luger!

@@

% describe saturation

% describe TV-min algorithm
We use the method of halo photometry, first described by \citet{White2017} and with further developments by \citet{Farr2018}.

% smear
This method differs from smear photometry \citep{Pope2016}, which uses collateral `smear' data to obtain a 1-D spatial profile with $\sim 1/1000$ of the flux on each CCD. While this has also been used for photometry of very bright stars, the `halo' method has the advantage of averaging over many more realizations of the K2 pointing systematics and permitting more significant calibration, achieving a generally higher photometric prevision.

@@

% incorporation of k2sc
While the TV-min procedure produces in most cases a fairly clean light curve, there are nevertheless residual systematic errors related to spacecraft motion. In order to correct these, we employ the \texttt{k2sc} code \citep{Aigrain2015,k2sc}, which simultaneously models a light curve as a 3D Gaussian Process (GP) in time and predicted position (the K2 standard data product \texttt{POS\_CORR}) in pixels $(x,y)$. The model prediction in time for fixed position is then a nonparametric model of the stellar variability, and the prediction in space evaluated for fixed time represents the pointing systematics and is subtracted from the input fluxes to obtain a final corrected flux, which is the time series we use and recommend for science.

@@

% short cadence availability 
Seven stars in Campaign~13 and one in Campaign~16 were not only assigned long-cadence halo apertures, but also downloaded at short cadence. TV-min halo photometry appears to perform poorly on short-cadence data, however, and we do not consider the data products for these stars to be satisfactory for an initial data release. This may be to do with the total variation being a poor proxy for noise in the light curves, with relatively fewer thruster-firing jumps and relatively higher white noise. The long-cadence data for these stars appear to be of normal quality.

@@

% do k2 smear after that
The sample of K2 bright stars presented here only includes those with halo apertures, but while some others are available conventionally, many were not assigned target pixels and were not downloaded at all. Smear photometry has been used to recover the brightest otherwise-unobserved stars in nominal \kepler (Pope et al., in prep.), and this can also be done in K2, although in the latter case the sample is much smaller due to competition with halo apertures and the systematics correction is more difficult. A natural extension of both pieces of work would be to produce smear light curves of all bright stars without halo apertures in K2, which would finally make the \kepler extended mission magnitude-complete at the bright end.

% tess halo 
The halo method naturally extends to other contexts where simple aperture photometry is not possible, such as for saturated stars observed by the Transiting Exoplanet Survey Satellite \citep[TESS;][]{tess}. Although the saturation limit is brighter ($T_{mag} \sim 6$) and this problem accordingly affects fewer stars and less badly, there are situations such as for $\alpha$~Centauri or $\beta$~Hydri where the bleed column reaches the edge of the chip and a SAP light curve is irrecoverable. We expect that TV-min halo photometry will therefore be valuable in ensuring that TESS can observe even the very brightest stars.

@@

BJSP acknowledges being on the traditional territory of the Lenape Nations and recognizes that Manhattan continues to be the home to many Algonkian peoples. We give blessings and thanks to the Lenape people and Lenape Nations in recognition that we are carrying out this work on their indigenous homelands. We would like to acknowledge the Gadigal Clan of the Eora Nation as the traditional owners of the land on which the University of Sydney is built and on which some of this work was carried out, and pay their respects to their knowledge, and to their elders past, present, and future.
%
